Systematic Design of Optimal Low-Thrust Transfers for the Three-Body Problem \thanks{This research has been supported in part by NSF under the grants CMMI-1243000 (transferred from 1029551), CMMI-1335008, and CNS-1337722.}

### 2017-07-27 Response.tex

\newcommand{\RNum}[1]{\uppercase\expandafter{\romannumeral #1\relax}}
\newcommand{\RI}{\text{\RNum{1}}}
\newcommand{\RII}{\text{\RNum{2}}}
\newcommand{\RIII}{\text{\RNum{3}}}

\newtheorem{definition}{Definition}
\newtheorem{lem}{Lemma}
\newtheorem{prop}{Proposition}
\newtheorem{cor}{Corollary}

\newenvironment{correction}{\begin{list}{}{\setlength{\leftmargin}{1cm}\setlength{\rightmargin}{1cm}}\vspace{\parsep}\item[]``}{''\end{list}}

\newcommand{\EditTL}[1]{{\color{red}\protect #1}}
%\renewcommand{\EditTL}[1]{{\protect #1}}

@@

(In the revised manuscript, the citation numbers for equations, assumptions, propositions, and references are changed. This answer is written according to the new item numbers.)

\subsection*{Reviewer 1} 
\begin{itshape}
Reviewer \#1: I have some issues with this paper that, in my opinion necessitate a major revision.
These issues are:
\end{itshape}

\begin{itemize}
    \item 
        \begin{itshape}
-It extensively re-covers issues that have been covered in many papers, making it difficult to disseminate what the original contributions are.  For example, I don't think the addition of the thrust term into the dynamics necessitates re-doing every piece of analysis that has been done with various resources in the literature about variational integrators (including comparisons with RK methods).  Things that can be found in other research should be cited, and summary comments can be used to point the readers to fill in this knowledge.  The content of this paper easily be reduced considerably to maintain focus on the original contributions.
    \end{itshape}

    Thank you for the comment, and we agree that there is repetition of material previously covered in other publications.
    The challenge is that the proposed approach utilizes several tools and concepts integrated together to design low-thrust orbital maneuvers in a systematic way, and in order to describe the key idea in a self-contained manner, each of prior approach should be described properly.

    To address this, the manuscript has been reduced and reorganized to enhance the clarity and clearly separate previously material and the contributions of this paper.
    The introduction has been reduced and the repetition and discuss on the variational integrator has been removed.
    The second section has been reorganized to ``Problem Formulation and Mathematical Background'' and has subsections which cover all of the background material, namely the planar circular restricted three body problem, Jacobi integral, invariant manifolds and \Poincare maps, the variational integrator. The section on variational integrators has been reduced and now only discusses the main developments of the integrator for the PCRTBP and omits the review of variational principles. The next sections introduce the optimal control formulation for the reachability analysis and the numerical examples, which are the main contributions of this paper.

    In short, the manuscript has been completely reorganized such that all of the background materials are summarized in Section 2. The remaining sections dealing with the proposed approach are expanded with more detailed descriptions and numerical results.

@@

\item 
    \begin{itshape}
The pros and cons of the PCRTBP

-The PCRTB gives rise to exploiting Poincare sections in a systematic manner.  2D solutions might be close enough for qualitative understanding of the solution space for many problems (as long as large plane changes aren't needed).

-The PCRTB is limiting for many pragmatic departure orbits (e.g. LEO and GTO, the most common departure orbits for low-thrust transfers).
\end{itshape}

Thank you for the comment and these characteristics of our choice of the PCRTBP. The authors would like to point out that the key idea of designing low-thrust orbits through the reachability analysis on the Poincar\'{e} section may be applied to other dynamic system. This manuscript illustrates such idea through PCRTB.

However, according to the reviewer's suggestion, additional details have been stated in the problem formulation section, including discussions on the choice of the PCRTBP and the relative strengths/weaknesses of this approach as copied below.

@@

\item 
    \begin{itshape}
The pros and cons of variational integrators should really be acknowledged rather than completely dismissing the RK methods which are still the current state of the art in mission design (using higher-orders and suitable tolerances)

-Pros = boundedness in the energy behavior.  Very long term integration will not deviate from the known dynamical structure.

-Cons = you don't know where you are on the energy surface, just that you are on it somewhere.  This is still a problem in actual trajectory design.  Knowing the error is on an energy surface does not mean that you don't have error.
\end{itshape}

Thank you for the comment and the detail on the benefits/drawbacks of variational integrators. The authors agree that the general-purpose numerical integrators such as higher-order RK methods are suitable in many cases. The proposed approach utilizes the variational integrator, particularly due to its numerical structural stability of capturing long-term behaviors of small effects accurately and efficiently, which is critical in low-thrust orbital maneuvers.

While the structural stability of geometric numerical integrators is often illustrated by the conserved quantities such as energy or momentum, it has been shown that variational integrators are also effective in computing the rate of energy change for non-conservative systems~[20].

To clarify these, additional discussion has been included in the paper in the subsection which introduces the variational integrator for the PCRTBP.
Much of the discussion of variational integrators in the introduction has also been removed.
The additional content, which discusses the relative benefits of variational integrators is copied below.

\begin{correction}
Both integration schemes result in trajectories that are initially nearly identical.
\Cref{fig:compare_energy} shows the mean Jacobi integral deviation over the entire simulation time as a function of computation time.
For a given computational effort, in the form of simulation run time, the variational integrator will provide a smaller energy deviation as compared to the \EditTL{conventional} integration scheme.
Over long simulation horizons or with the addition of small control inputs the inability of conventional integration schemes to accurately track the system energy limits the applicability of conventional techniques in which energy conservation is mandatory for characterizing the solution space.
In spite of this improved energy behavior, the order and design of the variational integrator still play an important role in the accuracy of the state vector.
This work uses a second order integrator which has improved energy behavior but potentially lower state accuracy as compared to high order Runge Kutta methods. 
However, for this preliminary analysis the variational integrator provides reduced computational demands and is useful to characterize and explore the trade space in preliminary mission design scenarios.
In addition, it has been shown that variable step integrators, such as \texttt{ODE45} in Matlab, tend to degrade first-order gradient based methods, such as those used in~\cref{sec:optimal_control}~\cite{pellegrini2016}.
The fixed step size of the presented variational integrator mitigates the \EditTL{artificial} ``noise'' of variable step integration schemes in gradient computation.
\EditTL{In summary, this paper utilizes the variational integrator to capture the long-term effects of low-thrust devices accurately and efficiently, and these numerical structural stability is critical for numerical optimization in the subsequent development.}
\end{correction}
\end{itemize}
From here on out \EditTL{our?} comments specific to various sections of the paper follow:

@@

\item 
    \begin{itshape}
"Since there is an insufficient number of analytical constants, or integral of motion, numerical methods must be used to investigate solutions to the three-body problem."  This isn't completely true as there are also approximations to the (uncontrolled) three-body motion, such such as the Richardson-Cary approximation.  Saying numerical methods are required may be a bit strong.  May want to say "often required".
\end{itshape}

Thank you for the comment.
The sentence has been modified as follows.

\begin{correction}
Finally, there is no closed form analytical solution for the three-body problem and there are an insufficient number of analytical constants~\cite{szebehely1967}.
As a result numerical methods are often required to investigate solutions to the three body problem.
\end{correction}

@@

\item 
    \begin{itshape}
"These techniques suffer from numerical instability and energy drift behaviors which make them ill-suited for long-term propagation. These dissipative effects are even more detrimental with the addition of low-thrust propulsion to the dynamic equations of motion."  It really depends on what you mean by "long term".  If transfers are on the order of hundreds of days, then high precision runge-kutte integrator, for example a Runge-Kutte Verner 8(9), can maintain suitable accuracy such that the energy drift in the Hamiltonian is very small, as can be seen in some of the references that the authors have cited.  If, on the other hand you mean 10-100's of years or more, i.e., when you create your Poincare maps, then this drift is more apparent.  I think that this statement needs to be quantified better because it gives a dishonest portrayal of high-precision Runge-Kutte integrators without numbers attached to it.
\end{itshape}

Thank you for the comment. Those numerical dissipations would depend on several parameters. But, for the particular numerical comparison illustrated by \Cref{fig:compare_energy}, such difference in conservation properties becomes noticeable in the simulation for 15 years. As suggested, the claims about the variational integrator have been qualified with some additional detail.
The correction is copied below.
\begin{correction}
As a result, these methods suffer from numerical instability and energy drift behaviors which make them ill-suited for the long-term propagation, e.g. on the order of decades to centuries, which is typical of the three-body problem. 
\end{correction}

\EditTL{Perhaps, the order of decades to centuries is too much. What is the typical orbital maneuver time for the numerical examples in Section 4?}

@@

\item
    \begin{itshape}
"Conventional integration techniques fail to capture the physical laws and geometric properties of the dynamic system [8]. As a result, the long term effects of low-thrust on the spacecraft trajectory are not accurately captured."  Again, I think situationally I would disagree.  Transfers on the order of 100's of days can maintain accuracies of 1e-10 (nondimensional units) in the Hamiltonian, which is generally considered a "small" amount of energy drift.
\end{itshape}

These sentences have been modified and combined to enhance clarity and reduce some of the unecessary bloat. 
In addition, some additional detail has been added to explicitly define the notion of ``long-term'' in the context of numerical integration. \EditTL{As shown by \Cref{fig:compare_energy}, the variational integrator exhibit smaller conservation errors than ODE45 at the similar computational efforts, for the particular numerical case considered. } The modifications are copied below.
\begin{correction}
    Conventional Runge-Kutta integration techniques, such as those implemented in~\cite{mingotti2011,grebow2011}, fail to capture the physical laws and geometric properties of the dynamic system.
    As a result, these methods suffer from numerical instability and energy drift behaviors which make them ill-suited for the long-term propagation, e.g. on the order of decades to centuries, which is typical of the three-body problem. 
\end{correction}

@@

\begin{correction}
While the \Poincare~section is well defined for the planar case, as a 2D representation, it is more challenging to implement and visualize for the general 3D three-body problem.
\end{correction}

\item
    \begin{itshape}
-"This discrete update map, or variational integrator, shares the same geometric properties of the continuous time system and exhibits much better energy behavior than the traditional integration methods, especially over long transfer durations with small magnitude control inputs."  Yes, but the order of the variational integrator still matters.  The drift in the state vector accuracy is still important to a mission designer, even if it is guaranteed to lie on an energy surface.  I think you should indicate that this is more for preliminary analysis to characterize the transfers, when this drift in state vector accuracy is tolerable.  I.E., it's useful for the introductory phases of mission design when trying to globally characterize the design space.  Characterizing the scope of the analysis in this way would be helpful to the audience.
\end{itshape}

Thank you for the comment.
The section on the variational integrator has been reduced and this specific sentence has been removed. 
However, additional detail has been provided in a the variational integrator subsection which discusses this drawback.

\begin{correction}
Over long simulation horizons or with the addition of small control inputs the inability of conventional integration schemes to accurately track the system energy limits the applicability of conventional techniques in which energy conservation is mandatory for characterizing the solution space.
In spite of this improved energy behavior, the order and design of the variational integrator still play an important role in the accuracy of the state vector.
This work uses a second order integrator which has improved energy behavior but potentially lower state accuracy as compared to high order Runge Kutta methods. 
However, for this preliminary analysis the variational integrator provides reduced computational demands, \EditTL{compared with general-purpose numerical integrators for the similar level of numerical errors}, and is useful to characterize and explore the trade space in preliminary mission design scenarios.
\end{correction}
\item 
    \begin{itshape}
"Optimal solutions are generally sensitive to small variations in the initial multipliers. As a result, the numerical stability of sensitivity derivatives is critical to accuracy and computational performance."  I don't agree that this implies that variational integrators are better for gradient-based methods.  The primary issue is that gradient-based methods that are solving optimal control problems are using first-order approximations to iterate on the problem.  The first-order approximation is what is poor, not necessarily the gradients.  Even if these gradients are of exact accuracy (and I agree that they should always be as accurate as possible), the first-order methods in the chaotic problem are what give the difficulties.  One could use a variational or standard integrator in a first-order gradient based method and nothing will change the fact that the method will be highly sensitive, save for partitioning the sensitivities as you did with your multiple shooting approach.  At best, you can change the statement to indicate that the accuracy in the gradients may improve, which is always helpful to gradient-based methods, but it's no guarantee that it eliminates major problem sensitivity.
\end{itshape}

It has been shown that the structure-preserving properties of geometric numerical integrators are also useful to numerical optimizations for improved computational efficiency and stability. For example, see the following reference.

\begin{itemize}
\item[-] Ober-Blöbaum, Sina, Oliver Junge, and Jerrold E. Marsden. "Discrete mechanics and optimal control: an analysis." ESAIM: Control, Optimisation and Calculus of Variations 17.2 (2011): 322-352.
\end{itemize}

The proposed approach utilizes desirable numerical properties of the variational integrator. However, the main contribution is systematic design of low-thrust orbital maneuvers through reachability set analysis on the Poincar\'{e} section. The above sentences may be confusing, and the authors decided to remove them from the introduction of the revised manuscript.

@@

\item
    \begin{itshape}
Authors may also want to point out the loss of accuracy in variable step integrators that can also degrade first-order gradient based methods.  I didn't see this anywhere and it's something that should be acknowledged.  Basically, variable step integrators will introduce artificial "noise" into the gradients:

Pellegrini, E.P., Russell, R.P., On the Accuracy of Trajectory State Transition Matrices, Paper AAS-15-785, AAS/AIAA Astrodynamics Specialist Conference, Vail, CO, Aug 2015
\end{itshape}

Thank you for the comment and helpful reference.

It is well known that variable step size may break conservational properties of a geometric numerical integrator. Variational integrators are no exception, and there has been a multisymplectic framework to address it. See

\begin{itemize}
\item[-] Lew, Adrian, et al. "Asynchronous variational integrators." Archive for Rational Mechanics and Analysis 167.2 (2003): 85-146.
\end{itemize}

However, the above formulation for structure-preserving, adaptive time-step variational integrators is beyond the scope of this paper, and we use a fixed time step throughout the presented numerical results.

Additional detail has been included in the manuscript discussing this potential drawback of variable step integrators.

\begin{correction}
In addition, it has been shown that variable step integrators, such as \texttt{ODE45} in Matlab, tend to degrade first-order gradient based methods, such as those used in~\cref{sec:optimal_control}~\cite{pellegrini2016}. \EditTL{The presented results are based on the variational integrator with a fixed time step for numerical simulation and gradient computation. }
\end{correction}

@@

\item 
    \begin{itshape}
"This results in an optimal solution rather than the suboptimal solution typical of direct optimal control methods." I disagree that suboptimality is typical of direct methods.  A direct method will typically solve the KKT conditions, which can be directly transcribed into the costates in the optimal control problem.  A direct method fully satisfying the KKT conditions can map to the indirect problem. If you want to strengthen this argument, then you should argue that it applies to direct methods which rely on coarse or suboptimal representations of the control.  For example, a hybrid direct method could use continuous optimal control theory control law, but directly minimize the cost function instead of solving the full TPBVP.  
\end{itshape}

Thank you for the comment.
The suboptimal claims of direct optimization has been reduced and modified.
Additional detail has been included on the merits of indirect optimal control vs. direct optimal control methods.
\begin{correction}
This results in an optimal solution rather than the suboptimal solution typical of direct optimal control methods which rely on \EditTL{coarse} or suboptimal control parameterizations.
For example, a hybrid direct method may use a continuous control law developed from indirect optimal control theory while directly minimizing a cost function instead of solving the full two-point boundary value problem~\cite{ozimek2010a}.
\end{correction}

@@

\item 
    \begin{itshape}
Section 5 - This method, while very useful in its given context is highly dependent on the use of the PCRTB in the framework of the current analysis.  This needs to somehow be conveyed.  3D transfers are significantly more challenging when it comes to exploiting Poincare sections.  It can be done, but more subplots, dimensions, and data storing are necessary to view the subspace - save to say, it is a more challenging endeavor that must be acknowledged for general purpose low-thrust mission design.
\end{itshape}

Thank you for the comment.
An additional paragraph has been added to the beginning of the mathematical background section which discusses our reachability based transfer approach.
This additional content is copied below.

\begin{correction}
The numerical examples presented in this section are designed in the context of the PCRTBP.
The dynamic environment has a four dimensional state space and offers a convient integration constant in the form of the Jacobi integral.
As a result, there are well defined methods to define and exploit \Poincare sections, which result in straightforward two-dimensional subspaces of the system.
Our approach uses the \Poincare section to approximate the reachability set on this reduced subspace.
As a result, this approach is more difficult to apply to three-dimensional transfers in the general three body problem.
\Poincare sections in the case of the general six dimensional state space are significantly more challenging and typically require more complicated visualization techniques. 
However, this is an area of active research and some of the authors future research is aimed at implementing this approach for non-planar transfer trajectories.
\end{correction}

\EditTL{Can we cite your recent AAS conference paper for orbital transfer around asteroids as a higher-order example?}

@@

There is additional research to extend these results to more general transfer scenarios in future work.
The incorporation of fourth body perturbations, such as the Sun in the Earth-Moon system, offers an additional method of increasing the reachable set with the combined use of the solar perturbation and low-thrust propulsion.
In addition, the assumed acceleration magnitude is currently beyond the capabilities of current electric propulsion systems.
Furthermore, this analysis did not consider the effect of variable mass on the optimal control solution.
This will result in a more complicated optimal control problem and is a focus of future research.
Finally, Lyapunov control theory, which has previously been applied to the two-body problem, is being investigated in the hope of designing closed loop control schemes for this three-body scenario~\cite{chang2002}.
The addition of attitude dynamics and realistic pointing constraints would significantly improve the applicability of this work.
\end{correction}

\item 
    \begin{itshape}
The authors should discuss some statistics of the convergence of their multiple shooting TPBVP - number of iterations, compute time, etc.
\end{itshape}

Thank you for the comment, however it seems to be rare for the research community to give detailed description of the numerical optimization method.
Most of the literature tends to not spend time discussing the low level details of the computational procedure, e.g.~\cite{grebow2011,stuart2010}.
In addition, the number of iterations, computation time and even the eventual cost function value are highly dependent on the specific system and more specifically the implementation of the algorithm in software. 
There are a wide variety of works which are focused more exclusively on the computational implementation of various optimal control techniques which do spend significant time on the computational characteristics.
However, our manuscript is primarily focused on demonstrating the feasbility of using reachability sets to design transfer trajectories.
As a result, there is not much benefit to detailed statistics of the implementation of this methodology, and we have omitted it from the manuscript.

### 2017-07-27 manuscript.tex
Incorporate some suggestions from Dr. Lee

There has been extensive research focused on the optimal control of spacecraft orbital transfers in the three-body problem~\cite{mingotti2011,grebow2011}. 
Inspite of the long history in astrodynamics, a number of difficulties arise when implenting optimal control techniques.
First, the optimal control problem is typically solved using direct methods, which approximate the continuous problem as a parameter optimization problem.
Alternatively, indirect methods apply the calculus of variations to derive the necessary conditions for optimality.
The indirect approach results in a lower dimensioned problem than the direct approach and provides algebraic conditions that guarantee local optimality in contrast to direct methods which generally rely on suboptimal representations of the control.
Second, in order to implement any optimization method a sufficiently accurate ``initial guess'' is required.
Frequently, insight into the problem or intuition on the part of the designer is often required to determine initial conditions that will converge to the desired solution due to the inherent nonlinear and chaotic behavior of the three-body system.
Finally, there is no closed form analytical solution for the three-body problem and there are an insufficient number of analytical constants~\cite{szebehely1967}.
As a result, numerical methods are often required to investigate solutions to the three body problem.
Conventional Runge-Kutta integration techniques, such as those implemented in~\cite{mingotti2011,grebow2011}, fail to capture the physical laws and geometric properties of the dynamic system.
Therefore, these methods suffer from numerical instability and energy drift behaviors which make them ill-suited for the long-term propagation, e.g. on the order of decades to centuries, which is typical of the three-body problem.

@@

In short, the authors present a systematic method of generating optimal transfer orbits in the three-body problem.
Previous results in the design of optimal transfers have relied on suboptimal direct optimization methods and conventional integration techniques.
This paper provides a discrete optimal control formulation to generate the reachability set on a \Poincare section.
We derive and solve the Euler-Lagrange equations to generate the necessary conditions for an optimal solution.
This results in an optimal solution rather than the suboptimal solution typical of direct optimal control methods which rely on course or suboptimal control parameterizations.
For example, a hybrid direct method may use a continuous control law developed from indirect optimal control theory while directly minimizing a cost function instead of solving the full two-point boundary value problem~\cite{ozimek2010a}.
In addition, the use of a geometric integrators ensures numerical stability for long-duration orbit transfers and maintains this behavior with the addition of small magnitude control inputs.
Our computation of the reachable set allows for a simple metric of defining optimal trajectories.
We avoid the issues inherent in selecting a valid initial condition for optimization.
Rather, we choose a state on the reachability set which minimizes the distance toward the desired target.
We demonstrate these capabilities via two numerical examples simulating transfer trajectories in the Earth-Moon system.

@@

A simulation comparing the variational integrator to a conventional Runge-Kutta method is presented in this section.
A particle is simulated from an initial condition of \( \vecbf{x}_0 = \begin{bmatrix} 0.75 & 0 & 0 & 0.2883\end{bmatrix}^T \) for \( t_f = 200 \approx 15\) years in the Earth-Moon system.
    The variational integrator uses a range of step sizes between \SIrange{4.72}{472}{\second} while the Runge-Kutta method uses a variable step size implemented via \texttt{ODE45} in Matlab.
    The step size of the variable integrator is varied to approximately match the run time required by the conventional \texttt{ODE45} integrator.
\Cref{fig:compare_trajectory} shows the trajectory of the spacecraft in the rotating reference for this comparison.
Both integration schemes result in trajectories that are initially nearly identical.
\Cref{fig:compare_energy} shows the mean Jacobi integral deviation over the entire simulation time as a function of computation time.
For a given computational effort, in the form of simulation run time, the variational integrator will provide a smaller energy deviation as compared to the convential integration scheme.
Over long simulation horizons or with the addition of small control inputs the inability of conventional integration schemes to accurately track the system energy limits the applicability of conventional techniques in which energy conservation is mandatory for characterizing the solution space.
In spite of this improved energy behavior, the order and design of the variational integrator still play an important role in the accuracy of the state vector.
This work uses a second order integrator which has improved energy behavior but potentially lower state accuracy as compared to high order Runge Kutta methods. 
However, for this preliminary analysis the variational integrator provides reduced computational demands and is useful to characterize and explore the trade space in preliminary mission design scenarios.
In addition, it has been shown that variable step integrators, such as \texttt{ODE45} in Matlab, tend to degrade first-order gradient based methods, such as those used in~\cref{sec:optimal_control}~\cite{pellegrini2016}.
The fixed step size of the presented variational integrator mitigates the artifical ``noise'' of variable step integration schemes in gradient computation.

@@

The numerical examples presented in this section are designed in the context of the PCRTBP.
The dynamic environment has a four dimensional state space and offers a convient integration constant in the form of the Jacobi integral.
As a result, there are well defined methods to define and exploit \Poincare sections, which result in straightforward two-dimensional subspaces of the system.
Our approach uses the \Poincare section to approximate the reachability set on this reduced subspace.
As a result, this approach is more difficult to apply to three-dimensional transfers in the general three body problem.
\Poincare sections in the case of the general six dimensional state space are significantly more challenging and typically require more complicated visualization techniques. 
However, this is an area of active research and some of the authors future research is aimed at implementing this approach for non-planar transfer trajectories.

@@

In~\cref{fig:multiple shooting}, we show a schematic representation of the multiple shooting procedure.
We split the optimal control horizon into equal length subsegments such that the length of each segment is \( \frac{k}{n} \), where \( k, n \) are the total number of steps and number of stages, respectively.
Similarly, we divide the state and costate trajectories into \( n \) equal segments. 
To ensure continuity, additional interior constraints are incorporated as
\begin{subequations}
\begin{align}\label{eq:interior_constraints}
        \vecbf{x}_1^{-} - \vecbf{x}_1^{+} &= 0 , \\ 
        \vecbf{\lambda}_1^{-} - \vecbf{\lambda}_1^{+} &= 0 , \\
        \vecbf{x}_2^{-} - \vecbf{x}_2^{+} &= 0 , \\ 
        \vecbf{\lambda}_2^{-} - \vecbf{\lambda}_2^{+} &= 0 , \\
        &\vdots \\
        \vecbf{x}_{n-1}^{-} - \vecbf{x}_{n-1}^{+} &= 0 , \\ 
        \vecbf{\lambda}_{n-1}^{-} - \vecbf{\lambda}_{n-1}^{+} &= 0.
\end{align}
\end{subequations}
Using the multiple shooting method reduces the sensitivity of the terminal states, \( \vecbf{x}_i^{+}, \vecbf{\lambda}_i^{+}\), to variations of the initial states, \(\vecbf{x}_{i-1}^{-}, \vecbf{\lambda}_{i-1}^{-}\).
As a result, the design vector \( \vecbf{\chi} \) is augmented with the additional interior initial conditions, \( \vecbf{x}_i^{-}, \vecbf{\lambda}^{-}\).
Similarly, the constraint vector is augmented with the additional interior constraints defined in~\cref{eq:interior_constraints}.
Based on experimentation, we use four stages in our multiple shooting method.
This provided the best performance and convergence stability while minimizing the difficulties in additional interior constraints.
The multiple shooting algorithm now varies the design vector \( \vecbf{\chi}\) to ensure that the constraints in \( \vecbf{G}(\vecbf{\chi}) \) are satisfied.
In this work, we use the Matlab nonlinear solver \texttt{fsolve} to solve the system of nonlinear equations defined by the multiple shooting algorithm.
Within \texttt{fsolve}, we use the trust-region dogleg solver which makes use of the Powell dogleg procedure for computing a step direction and magnitude to minimize successive iterations of the solver.
All numerical integration is performed using the discrete variational integrator described in~\cref{sec:discrete_var}.
% tpbvp is solved using a shooting method (fsolve in matlab)
% determine a design vector X of variables we need to determine
% define a constraint vector G(X) that we need to satisfy
% multiple shooting has interior points of continuity constraints
% sensitivity of terminal state is used to correct the initial condition (cite a book/paper)
% plot of multiple shooting ( stages)
\section{Numerical Example}\label{sec:simulation}
We present two numerical simulations in the Earth-Moon system to demonstrate the transfer procedure.
These simulations enable the spacecraft to depart from the natural dynamics through the use of low-thrust propulsion.
The reachability set on the \Poincare section allows for a straightforward method of determining transfer opportunities.
The first example is a transfer from a periodic orbit of the \( L_1 \) Lagrange point to a fixed orbit of the moon.
This example uses a single iteration of the reachable set computation in the design of the transfer.
The second example is a transfer from geostationary orbit of the Earth to a period orbit of \( L_1 \). 
This examples demonstrates the ability to extend the reachability process to multiple iterations, to allow for a much larger and more general transfer.
With both examples it is possible to depart from the vicinity of the Earth to a Moon orbit via a series of reachable sets defined on \Poincare sections.

>>>>

There has been extensive research focused on the optimal control of spacecraft orbital transfers in the three-body problem~\cite{mingotti2011,grebow2011}. 
Inspite of the long history in astrodynamics, a number of difficulties arise when implenting optimal control techniques.
First, the optimal control problem is typically solved using direct methods, which approximate the continuous problem as a parameter optimization problem.
Alternatively, indirect methods apply the calculus of variations to derive the necessary conditions for optimality.
The indirect approach results in a lower dimensioned problem than the direct approach and provides algebraic conditions that guarantee local optimality in contrast to direct methods which generally rely on suboptimal representations of the control.
Second, in order to implement any optimization method a sufficiently accurate ``initial guess'' is required.
Frequently, insight into the problem or intuition on the part of the designer is often required to determine initial conditions that will converge to the desired solution due to the inherent nonlinear and chaotic behavior of the three-body system.
Finally, there is no closed form analytical solution for the three-body problem and there are an insufficient number of analytical constants~\cite{szebehely1967}.
As a result, numerical methods are often required to investigate solutions to the three body problem.
Conventional Runge-Kutta integration techniques, such as those implemented in~\cite{mingotti2011,grebow2011}, fail to capture the physical laws and geometric properties of the dynamic system.
Therefore, these methods suffer from numerical instability and energy drift behaviors which make them ill-suited for the long-term propagation, e.g. on the order of decades to centuries, which is typical of the three-body problem. 
Furthermore, the variational integrator can allow for reduced computational effort without any loss of numerical stability or energy drift.

@@

In short, the authors present a systematic method of generating optimal transfer orbits in the three-body problem.
Previous results in the design of optimal transfers have relied on suboptimal direct optimization methods and conventional integration techniques.
This paper provides a discrete optimal control formulation to generate the reachability set on a \Poincare section.
We derive and solve the Euler-Lagrange equations to generate the necessary conditions for an optimal solution.
This results in an optimal solution rather than the suboptimal solution typical of direct optimal control methods which rely on coarse or suboptimal control parameterizations.
For example, a hybrid direct method may use a continuous control law developed from indirect optimal control theory while directly minimizing a cost function instead of solving the full two-point boundary value problem~\cite{ozimek2010a}.
In addition, the use of a geometric integrators ensures numerical stability for long-duration orbit transfers and maintains this behavior with the addition of small magnitude control inputs.
Our computation of the reachable set allows for a simple metric of defining optimal trajectories.
We avoid the issues inherent in selecting a valid initial condition for optimization.
Rather, we choose a state on the reachability set which minimizes the distance toward the desired target.
We demonstrate these capabilities via two numerical examples simulating transfer trajectories in the Earth-Moon system.

@@

A simulation comparing the variational integrator to a conventional Runge-Kutta method is presented in this section.
A particle is simulated from an initial condition of \( \vecbf{x}_0 = \begin{bmatrix} 0.75 & 0 & 0 & 0.2883\end{bmatrix}^T \) for \( t_f = 200 \approx 15\) years in the Earth-Moon system.
    The variational integrator uses a range of step sizes between \SIrange{4.72}{472}{\second} while the Runge-Kutta method uses a variable step size implemented via \texttt{ODE45} in Matlab.
    The step size of the variable integrator is varied to approximately match the run time required by the conventional \texttt{ODE45} integrator.
\Cref{fig:compare_trajectory} shows the trajectory of the spacecraft in the rotating reference for this comparison.
Both integration schemes result in trajectories that are initially nearly identical.
\Cref{fig:compare_energy} shows the mean Jacobi integral deviation over the entire simulation time as a function of computation time.
For a given computational effort, in the form of simulation run time, the variational integrator will provide a smaller energy deviation as compared to the conventional integration scheme.
Over long simulation horizons or with the addition of small control inputs the inability of conventional integration schemes to accurately track the system energy limits the applicability of conventional techniques in which energy conservation is mandatory for characterizing the solution space.
In spite of this improved energy behavior, the order and design of the variational integrator still play an important role in the accuracy of the state vector.
This work uses a second order integrator which has improved energy behavior but potentially lower state accuracy as compared to high order Runge Kutta methods. 
However, for this preliminary analysis the variational integrator provides reduced computational demands, as compared with general-purpose numerical integrators with a similar level of numerical error, and is useful to characterize and explore the trade space in preliminary mission design scenarios.
In addition, it has been shown that variable step integrators, such as \texttt{ODE45} in Matlab, tend to degrade first-order gradient based methods, such as those used in~\cref{sec:optimal_control}~\cite{pellegrini2016}.
The presented results are based on the variational integrator with a fixed time step for numerical simulation and gradient computation.
The fixed step size of the presented variational integrator mitigates the artificial ``noise'' of variable step integration schemes in gradient computation.
In summary, this paper utilizes the variational integrator to capture the long-term effects of low-thrust devices accurately and efficiently.
The structural stability during numerical integration is critical for numerical optimization in the subsequent developments.

@@

The numerical examples presented in this section are designed in the context of the PCRTBP.
The dynamic environment has a four dimensional state space and offers a convient integration constant in the form of the Jacobi integral.
As a result, there are well defined methods to define and exploit \Poincare sections, which result in straightforward two-dimensional subspaces of the system.
Our approach uses the \Poincare section to approximate the reachability set on this reduced subspace.
As a result, this approach is more difficult to apply to three-dimensional transfers in the general three body problem.
\Poincare sections in the case of the general six dimensional state space are significantly more challenging and typically require more complicated visualization techniques. 
However, this is an area of active research and some of the authors future research is aimed at implementing this approach for non-planar transfer trajectories~\cite{kulumani2016d}.

@@

In~\cref{fig:multiple shooting}, we show a schematic representation of the multiple shooting procedure.
We split the optimal control horizon into equal length subsegments such that the length of each segment is \( \frac{k}{n} \), where \( k, n \) are the total number of steps and number of stages, respectively.
Similarly, we divide the state and costate trajectories into \( n \) equal segments. 
To ensure continuity, additional interior constraints are incorporated as
\begin{subequations}
\begin{align}\label{eq:interior_constraints}
        \vecbf{x}_1^{-} - \vecbf{x}_1^{+} &= 0 , \\ 
        \vecbf{\lambda}_1^{-} - \vecbf{\lambda}_1^{+} &= 0 , \\
        \vecbf{x}_2^{-} - \vecbf{x}_2^{+} &= 0 , \\ 
        \vecbf{\lambda}_2^{-} - \vecbf{\lambda}_2^{+} &= 0 , \\
        &\vdots \\
        \vecbf{x}_{n-1}^{-} - \vecbf{x}_{n-1}^{+} &= 0 , \\ 
        \vecbf{\lambda}_{n-1}^{-} - \vecbf{\lambda}_{n-1}^{+} &= 0.
\end{align}
\end{subequations}
Using the multiple shooting method reduces the sensitivity of the terminal states, \( \vecbf{x}_i^{+}, \vecbf{\lambda}_i^{+}\), to variations of the initial states, \(\vecbf{x}_{i-1}^{-}, \vecbf{\lambda}_{i-1}^{-}\).
As a result, the design vector \( \vecbf{\chi} \) is augmented with the additional interior initial conditions, \( \vecbf{x}_i^{-}, \vecbf{\lambda}^{-}\).
Similarly, the constraint vector is augmented with the additional interior constraints defined in~\cref{eq:interior_constraints}.
Based on experimentation, we use four stages in our multiple shooting method.
This provided the best performance and convergence stability while minimizing the difficulties in additional interior constraints.
The multiple shooting algorithm now varies the design vector \( \vecbf{\chi}\) to ensure that the constraints in \( \vecbf{G}(\vecbf{\chi}) \) are satisfied.
In this work, we use the Matlab nonlinear solver \texttt{fsolve} to solve the system of nonlinear equations defined by the multiple shooting algorithm.
Within \texttt{fsolve}, we use the trust-region dogleg solver which makes use of the Powell dogleg procedure for computing a step direction and magnitude to minimize successive iterations of the solver.
All numerical integration is performed using the discrete variational integrator described in~\cref{sec:discrete_var}.
% tpbvp is solved using a shooting method (fsolve in matlab)
% determine a design vector X of variables we need to determine
% define a constraint vector G(X) that we need to satisfy
% multiple shooting has interior points of continuity constraints
% sensitivity of terminal state is used to correct the initial condition (cite a book/paper)
% plot of multiple shooting ( stages)
\section{Numerical Example}\label{sec:simulation}
We present two numerical simulations in the Earth-Moon system to demonstrate the transfer procedure.
These simulations enable the spacecraft to depart from the natural dynamics through the use of low-thrust propulsion.
The reachability set on the \Poincare section allows for a straightforward method of determining transfer opportunities.
The first example is a transfer from a periodic orbit of the \( L_1 \) Lagrange point to a fixed orbit of the moon.
This example uses a single iteration of the reachable set computation in the design of the transfer.
The second example is a transfer from geostationary orbit of the Earth to a period orbit of \( L_1 \). 
This examples demonstrates the ability to extend the reachability process to multiple iterations, to allow for a much larger and more general transfer.
With both examples it is possible to depart from the vicinity of the Earth to a Moon orbit via a series of reachable sets defined on \Poincare sections.
The numerical examples presented in this section satisfy the necessary conditions for local optimality and obtaining a globally optimal solution is considered beyond the schope of this paper.
